---
title: "lab_02"
author: "Derek Willis, adapted by Aidan Hughes"
date: "2024-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

For this lab we'll be working with a slice of Maryland voter registration data: records of people who have registration dates this year through Aug. 7. Let's load it.

```{r}
md_voters_2024 <- read_rds("data/lab_02.rds")
```

## Examine Data

Take a look at the data, getting familiar with what it contains. Most of the columns should be self-explanatory, but there are two date columns: STATE_REGISTRATION_DATE and COUNTY_REGISTRATION_DATE. The former represents the date the person registered as a Maryland voter; the latter is when they registered in their current county. They may be the same, but if a voter moves from one county to another they will differ. This is important because if the two dates are the same, that's a pretty good indication that this is a *new* voter, someone who has no history of registering in Maryland. You'll need to keep that definition in mind.

## Answer questions

**Q1** How many voters appear to be *new* registrants since July 21, when President Joe Biden announced he would leave the race, based on the information above? Create a dataframe to hold the only those voters.

Which counties have seen the most? Take a look at county population figures here and compare that list with your data findings in your answer: https://worldpopulationreview.com/us-counties/maryland.

**A1**

```{r}
new_registration <- md_voters_2024 |>
filter (COUNTY_REGISTRATION_DATE == STATE_REGISTRATION_DATE) |>
filter (STATE_REGISTRATION_DATE >= "2024-07-21")
nrow(new_registration)

```

**Q2** A better question is whether the new voters you found in Q1 are an outlier compared to other weeks this year. Using the original dataframe with 2024 registrants, use `mutate()` and lubridate to produce week numbers from the COUNTY_REGISTRATION_DATE (Lubridate has a function that turns a date into a week number. Use that to answer this question). Save that to the original dataframe. Then calculate how many voters are in each week. Describe the distribution of voters groups in a sentence - are the weeks from July 21-August 7 different from the rest, and if so, how?

**A2**
Weeks of July 21 to August 7 is numeric week 30 and 31. These two weeks are towards the top of the new voters count list as they are 2 of three weeks with over 7000 new registered voters. Around this time seems to be the peak of new voters with the other weeks farther from these dates seeming to decrease in new registered voters. 
```{r}
md_voters_2024 |>
  select(COUNTY_REGISTRATION_DATE) |>
     mutate(
       week_num = week(COUNTY_REGISTRATION_DATE)) |>
        group_by(week_num) |> 
        summarise(count_voters = n()) |>
        arrange(desc(count_voters))
```

**Q3** Using lubridate, show which month saw the largest number of registrations. Does the answer help shed more light on Q2?
**A3** 

```{r}
md_voters_2024 |>
  select(COUNTY_REGISTRATION_DATE) |>
     mutate (month = floor_date(COUNTY_REGISTRATION_DATE, "month")) |>
      group_by(month) |>
        summarise(count_voters_m = n()) |>
        arrange(desc(count_voters_m))
```

**Q4** Maryland has, officially, five regions: https://www.visitmaryland.org/article/maryland-regions. Modify the original dataframe again to add a `region` column based on the `COUNTY` value using case_when. Are you going to have to do this for every Maryland county plus Baltimore City? Yes, you are.

Which regions and parties account for the most new voters?

**A4** 

```{r}
maryland_regions <- md_voters_2024 |>
  mutate(
    region = case_when(
      
      COUNTY == "Garret" ~ "Western Region",
      COUNTY == "Allegany" ~ "Western Region",
      COUNTY == "Washington" ~ "Western Region",
      COUNTY == "Frederick" ~ "Capital Region",
      COUNTY == "Montgomery" ~ "Capital Region",
      COUNTY == "Prince George's" ~ "Capital Region",
      COUNTY == "Anne Arundel" ~ "Central Region",
      COUNTY == "Baltimore City" ~ "Central Region",
      COUNTY == "Baltimore County" ~ "Central Region",
      COUNTY == "Carroll"~ "Central Region",
      COUNTY == "Howard" ~ "Central Region",
      COUNTY == "Calvert"  ~ "Southern Region",
      COUNTY == "Charles" ~ "Southern Region",
      COUNTY == "St. Mary's"  ~ "Southern Region",
      COUNTY == "Kent" ~ "Eastern Shore Region",
      COUNTY == "Queen Anne's" ~ "Eastern Shore Region",
      COUNTY == "Talbot" ~ "Eastern Shore Region",
      COUNTY == "Caroline" ~ "Eastern Shore Region",
      COUNTY == "Dorchester" ~ "Eastern Shore Region",
      COUNTY == "Wicomico" ~ "Eastern Shore Region",
      COUNTY == "Somerset" ~ "Eastern Shore Region",
      COUNTY == "Worcester" ~ "Eastern Shore Region",
      COUNTY == "Cecil" ~ "Eastern Shore Region"
      )
  )
  head(maryland_regions)
  view(maryland_regions)
```


**Q5** Filter md_voters_2024 to only include voters from the Capital region or the Western region. Save the data in a new variable/data set named capital_or_western.

How many voters are in the new data set?

48559

```{r}
capital_or_western <- maryland_regions |>

  filter(region == "Capital Region" | region == "Western Region")

nrow(capital_or_western)
```


**Q6** Finally, filter md_voters_2024 to only include voters who meet BOTH of these conditions:

1) they are from the Eastern Shore region or Western region
2) their party affiliation is DEM and their Gender is F

Hint: you can accomplish this with two filters

Save the data in a new variable/data set named final_filter.

How many voters are in the new data set?
12279

```{r}
capital_or_western <- maryland_regions |>

  filter(region == "Capital Region" | region == "Western Region") |>
  filter(PARTY == "DEM" & GENDER == "F")

nrow(capital_or_western)
```




**BONUS QUESTION** What's an additional journalistic question you could ask of this data? Show me the code you'd use to do it below, or at least describe the logic step-by-step in English if you're not sure how to write the code for it yet.


**A** 
I could also filter to see how many new registered voters from July 21 didn't register with any specific party. 
```{r}
new_unaffiliated <- md_voters_2024 |>
filter (COUNTY_REGISTRATION_DATE == STATE_REGISTRATION_DATE) |>
filter (STATE_REGISTRATION_DATE >= "2024-07-21") |>
filter (PARTY == "UNA")
nrow(new_unaffiliated)
```

