---
title: "lab_03"
author: "Derek Willis, adapted by Aidan Hughes"
date: "2024-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)

# Load janitor
library(janitor)
```

### Load data

```{r}
# Load MD slice of FEC House candidate contribution data
md_contribs <- read_rds("data/lab_03.rds")
```

### Do this first: Fix header issues

You'll need to fix one issue with the column names using janitor. Write a code block to do that, making sure that you save the results to the same dataframe.

```{r}

cleaned_md_contribs <- md_contribs |>
  clean_names()

cleaned_md_contribs

```


### **Q1**: 
Which seven cities in Maryland had the largest *number* of contributions from donors? Bethesda has the largest number of contributions 

How many did each have? 
They had 1,177 total contributions

To answer this question, you'll have to clean the messy "contributor_city" field.

Note: To answer this question, you'll need to see every variation of the "contributor_city" for the top seven cities, which you should identify through some initial analysis. Then find variations of them and change those variations into the most common version.

Hint: you should be starting out with a group_by and summarize just like we did in week 1 and 2. That'll help you identify the top seven names, and then you can click through the data to find variations of those names.
**Top 7 ORIGINAL**
BETHESDA 1177			
CHEVY CHASE	704			
BALTIMORE	690			
ANNAPOLIS	490			
POTOMAC	490			
SILVER SPRING	405			
ROCKVILLE	358

**GOOGLE USED TO LOOK UP IF BETHESDA AND NORTH BETHESDA WAS THE SAME -- GOOGLE SAID THEY WERE DIFFERENT -- SAME THING FOR POTOMAC AND NORTH POTOMAC -- AND ANNAPOLIS AND ANNAPOLIS JUNCTION**

Then, look back at Task 15 in Pre-Lab 3 for a clue about how to update specific values. Don't forget, you'll need a step at the end to ensure that everything aside from the values you're changing stay as the default value.

**A1**:
Which seven cities in Maryland had the largest *number* of contributions from donors? Bethesda has the largest number of contributions 

How many did each have? 
They had 1,177 total contributions
```{r}
md_contribs |>
select(contributor_city) |>
 group_by(contributor_city) |>
  summarise( 
    count_contribs = n()
  )|>
  arrange(desc(count_contribs))


cleaned_md_contribs <- md_contribs |>
  clean_names() |> 
    distinct() |>
    mutate(contributor_city = str_to_title(contributor_city)) |>
    mutate(contributor_city = case_when(
    contributor_city == "POTAMIC" ~ "POTOMAC",
    contributor_city == "Potamic" ~ "POTOMAC",
    contributor_city == "ROCKWELL" ~ "ROCKVILLE",
    contributor_city == "ROCK HALL" ~ "ROCKVILLE",
    contributor_city == "BLATIMORE" ~ "BALTIMORE",
    contributor_city == "BALRIMORE" ~ "BALTIMORE",
    contributor_city == "ANAPOLIS" ~ "ANNAPOLIS",
    contributor_city == "SILVER SPRINGS" ~ "SILVER SPRING",
    TRUE ~ contributor_city))

cleaned_md_contribs


```

### **Q2:** Find the 5-digit ZIP code in Maryland that had the highest total amount of contributions (by dollar value).

Where is that zip code and what kind of area is it, demographically? 20817, This is the Bethesda zip code and Bethesda is a very rich upperclass so it makes sense that they had a high amount of total dollars contributed

Which committee is the largest recipient of money from that zip code? 
		YVETTE4CONGRESS, ADAM GRAY FOR CONGRESS	
C00875336	BILL HAMLEN FOR CONGRESS  had the largest money that zip code at 3300

How many of the top five recipients are Maryland candidates?
Only one of them DAN COX FOR CONGRESS, INC.

Hint: You will need to use mutate and str_sub to ensure that you have a zip code column with the first five characters from the existing zip code column.

Other functions you'll need afterwards are filter, group by, summarise, and arrange.

Break this down into individual steps and check your work along the way. That'll make this question more manageable.

**A2:**

```{r}
zip_cleaned_md_contribs <- cleaned_md_contribs |>
  distinct() |>
  mutate(contributor_zip = str_sub(contributor_zip, start=1, end=5)) |>
  group_by(contributor_zip) |>
  summarise(total_amount = sum(contribution_receipt_amount)) |>
  arrange(desc(total_amount))
  zip_cleaned_md_contribs
  
  
 largest_20817_recipient <- cleaned_md_contribs |>
   filter(contributor_zip == "20817") 
  head(largest_20817_recipient)
  

top_20817_committee <- largest_20817_recipient |>
  select(committee_id, committee_name, contribution_receipt_amount, contributor_zip, contributor_city, contributor_state) |>
  arrange(desc(contribution_receipt_amount))
 head(top_20817_committee)
  
```

### **Q3:**

Let's look at top individual donors. 
Write code to generate a list of top donors in each city, adding up the total amount of contributions. 
What does the original data say about the top two donors? 
One of the donors is from Rockville and the other one is from baltimore
Who are they and what do they do? 
KRACOV, DANIEL _ He is a partner and co-chair of clobal life sciences and healthcare regulatory pracitce in DC. 
ALLEN, MARGARET - she is the CEO of AGM Financial Services
Do you see anything in the list of top 10 donors that would make you uncertain about the accuracy of these results?
There is a Mark and Melonie Jacobsen both from Bethesda which could be odd in the sense that they donated separately and different number of contributions. 

Hint: you'll need to put two columns in a group_by to answer this question. What are the two categories you're trying to group the data by? Look back at Task 8 in Pre-lab 1 if you need a reminder of how to do this.

**A3:**

```{r}
top_donors_city <- cleaned_md_contribs |> 
  filter(contributor_city == "Baltimore") |>
  filter(contributor_city == "Annapolis") |>
  filter(contributor_city == "Potomac") |>
  filter(contributor_city == "Rockville") |>
  filter(contributor_city == "Silver Spring") |>
  filter(contributor_city == "Chevy Chase") |>
  filter(contributor_city == "Bethesda")|>
  summarise(
    count_contribs = n()
  ) |>
 arrange(desc(count_contribs))
 
head(top_donors_city)



top_donors <- cleaned_md_contribs |> 
  group_by(contributor_name, contributor_city) |>
  summarise(
    count_contribs = n()
  ) |>
  arrange(desc(count_contribs))

head(top_donors, 10)
```


### **BONUS QUESTION**

### **Q:** What's another problem you can identify in the data? How would it impact your ability to do an accurate analysis of something? (Three bonus points)

### Then, write code that would fix the issue. (Five bonus points)

**A:**

```{r}

```

```{r}

```



