---
title: "lab_05"
author: "derek willis, updated by Aidan Hughes"
date: "2026-10-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# load the tidyverse
install.packages('refinr')
options(scipen=999)
library(dplyr)
library(tidyverse)
library(lubridate)
library(refinr)
```

## Getting Started

This week we'll be combining and joining data to make it more useful and to ask some questions about it. We'll revisit our new Maryland voters data and try to join it with contributions to the Harris campaign from Maryland donors made via ActBlue. Can we find new voters who are not registered Democrats who gave to Harris after she became a presidential candidate on July 21? What, if anything, can we say about those people?

We'll be doing something similar to what Ence Morse did for [this Washington Post story](https://www.washingtonpost.com/politics/2024/09/24/kamala-harris-younger-donors/?pwapi_token=eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJyZWFzb24iOiJnaWZ0IiwibmJmIjoxNzI3MzIzMjAwLCJpc3MiOiJzdWJzY3JpcHRpb25zIiwiZXhwIjoxNzI4NzA1NTk5LCJpYXQiOjE3MjczMjMyMDAsImp0aSI6ImIwYWMxNTk4LWMwMDgtNDE5Ny04MjAyLWVmMWNjZWFmYWZiMiIsInVybCI6Imh0dHBzOi8vd3d3Lndhc2hpbmd0b25wb3N0LmNvbS9wb2xpdGljcy8yMDI0LzA5LzI0L2thbWFsYS1oYXJyaXMteW91bmdlci1kb25vcnMvIn0.3-s0tBkMV7Y9DD21CBsKMlaliGlOKT2egCO-UFT735Y), just on a smaller scale.

## Tasks

**First, load two CSV contribution files in the `data` folder, one each for donations made in July and August, and then combine them into a single dataframe.**

```{r}
july_data_harris <- read_csv("data/july_harris.csv")
august_data_harris <- read_csv("data/august_harris.csv")


july_august_combined <- bind_rows(july_data_harris, august_data_harris)
View(july_august_combined)


```


**Next, let's read in the new voters (people who registered in 2024):**

```{r}
new_voters <- read_csv("data/new_voters.csv")
View(new_voters)
```

**Next join the Harris donations to the voter data. Remember, people's names are common, __so you'll want to use name columns and address information__. You'll need to figure out which columns to use from both dataframes.**

The new_voters data has both a single Address column and various parts of the address, such as HouseNumber. Make a plan for picking your join columns - try out some combinations - and save the results to a new dataframe.

Instead of a left_join, use the function inner_join this time. Unlike left_join or right_join, inner_join gets rid of any rows in BOTH data sets if there wasn't a match.

Using multiple columns in your join_by just requires you to separate them with commas. It should look like this:

join_by(column1,
        column2,
        column3)

Note that you might get a "many to many" error when you run the code, depending on the columns you pick That's R letting you know that something in one data set matched with multiple rows in the other data set. That's something really important to keep your eye out for, since R will perform the match against all those rows, meaning you might wind up with multiple extra rows when you should only have one.

You'll want to minimize the amount of rows that this is happening to by picking smart join columns, but we probably won't eliminate it completely. We won't deal with that for the purpose of this lab, but it's important to be aware of.

```{r}

new_voters_july_august_harris <- new_voters|> 
  inner_join(july_august_combined,
  join_by(FirstName == 'first_name', LastName == 'last_name', Address == 'address_one'))
  
  View(new_voters_july_august_harris)

# join the combined Harris file to the new voters
# you will need to determine what columns to use
```

## Answer questions


**Q1. Explain why you used the columns you did to make the join, and why that result is better than other ways you tried. Are there any problems with the join you made, such as the possibility of false positives?**


A1. 
The columns I used to make a join were the first and last name columns along with the persons address, because those are unique features that would be best able to identify any similarties when combining the data together. It is pretty unlikely that there would be more than one person that had the same, first and last name along with the same address. When I tried to join the data I did get a warning message that said that two rows had matched and could be false positives, it were rows 104 and 64775.

**Q2. Based on your join, how many donations to Harris have come from people who registered this year? How many unique people does that represent, and which donors have given the largest _number_ of donations since July 20?** 
1. 365 - all of them 
2. It represents 365 unique people 
3. SOPHIA HOFFER-PERKINS, BRITTANY MUSOLINO, and ERIC SNIPE all donated 7 times
A2.


```{r}
harris_donations_2024 <-new_voters_july_august_harris |>
  filter(CountyRegistrationDate >= '2024-01-01')|>
nrow(harris_donations_2024)
view(harris_donations_2024)
```
```{r}

largest_july_20 <-new_voters_july_august_harris |>
  filter(date >= '2024-07-20') |>
  rename(last_name = LastName, first_name = FirstName) |>
  group_by(last_name, first_name) |>
  summarise(
    count_contribs = n()
  ) |>
  arrange(desc(count_contribs))
head(largest_july_20)
 

```

**Q3. What are the party and gender breakdowns of those donors? How many donations has Harris gotten from non-Democrats?**


A3.
1. 190 Female donors, 163 Male donors, and 12 donors that didn't list Male or Female
2. 43 donations from non democrats
(I looked up how do use the does not equal because I forgot for the non dems section)



```{r}
gender_breakdown <- new_voters_july_august_harris |>
  group_by(Gender) |> 
  summarise(
    total= n()
  )
head(gender_breakdown)
```

```{r}
party_breakdown <- new_voters_july_august_harris |>
  group_by(Party) |> 
  summarise(
    total = n()
  )
head(party_breakdown)

non_dem_donation <- new_voters_july_august_harris |>
  filter(Party != "DEM")
nrow(non_dem_donation)
```
**Q4. Choose your own adventure: I want you to come up with a question you could ask of one of the dataframes we've created for this lab. It could involve just the new voters, or the donors or the joined dataframe, but it cannot be a variation of a previous question (so, for example, you can't ask, "How many Republicans have given to Harris").**

**First write the question and then try to answer it by writing code. If you're not sure of the exact code, try googling (or ask me for advice) but, at a minimum, describe what you think you'd have to do to answer it and what the results would look like. That can help.**

My question is how does the average donation amoung differ between male and female donors across the different political groups Democrats, Republicans and Unaffiliated/Independent Voters registered in 2024? - I am asking this question because typically woman are more likely to donate to political campaigns, but Men are more likely to give larger amounts of money to political campaigns

A4:
```{r}
donation_amount_gender_party <-new_voters_july_august_harris |>
  group_by(Gender, Party) |>
  summarise(
    total_donations = n(), 
    average_donation = mean(amount),
    median_amount = median(amount)
    ) |>
  arrange(Party, Gender)
head(donation_amount_gender_party)
```

**BONUS QUESTION: What's another publicly available data set you could join with these to ask additional questions? You'll need to provide a link to the data, what question(s) it would allow you to ask, and briefly describe any issues that might prevent you from joining the data without additional cleaning or manipulation (2 points).**
https://campaignfinance.maryland.gov/public/cf/downloads
I accessed the Maryland Financies from the Maryland State Board of Elections for 2025. My question is how donations especially for the democratic party changed, the following year after harris wasn't elected. To see if there was more money donated to the democrats by democrats.

There did say there was a warning about parsing issues so I did have to find the problem for the data. There says that the expected was 2 columns but ended up with 28 columns. I don't really know how to solve the problem cause I can
t even load it to see it. the glimpse financials wont show the whole data set even with the errors. 


**Then, download the data and add it to the "data" folder of this lab. Write the code needed to perform the join. (4 points)**

I did try to join them anyway and it showed up as an error. 

```{r}
md_financials_2025 <- read_csv("data/maryland_contributions_2025.csv")
problems(md_financials_2025)
md_financials_2025 <- read_csv("data/maryland_contributions_2025.csv", guess_max=177010)
glimpse(md_financials_2025)

md_financials_2025_harris <- bind_rows(july_data_harris, august_data_harris, md_financials_2025)
View(md_financials_2025_harris)
```



