---
title: "lab_01"
author: "Derek Willis, adapted by Aidan Hughes"
date: "2026-08-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of WinRed contribution data from West Virginia, which is in a file called `lab_01.rds`

**Task** Using the example from the pre_lab, write code to load the RDS file and assign it to a good variable name.

```{r} 
westvirginia_winred_contributions <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.

```{r}
glimpse(westvirginia_winred_contributions)
```


**Q1** How many rows and columns does this dataframe have?

**A1**
There are 24 columns and 40,441 rows 

## Answer questions

**Q2** Which West Virginia city provided the largest number of contributions? How many?
Charleston had the largest number of contributions, they had 3226 contributions.  

Hint: you'll need to use group_by() |> summarise() here. What column are we grouping by? Then, what should go inside the summarise() part? Then, rather than clicking through the 500+ cities to find the largest one, use the same function we used in our pre-lab to sort the data from largest number to the smallest.


**A2**
The column we are grouping by is the city and what should go inside the summarise is to count the contributions then arrange the contribution counts in decending order. 
```{r}
westvirginia_winred_contributions |>
 group_by(city) |>
  summarise( 
    count_contribs = n()
  )|>
  arrange(desc(count_contribs))
```

**Q3** Which West Virginia cities accounted for the fewest contributions? You don't have to name all of them, but list a few. How many of those cities are there, roughly speaking?

**A3**
Allen Junction, Alum Bridge, Bridgeport and Bergoo all show to be one of the cities with the the fewest contributions at one contribution. 
```{r}
westvirginia_winred_contributions |>
 group_by(city) |>
  summarise( 
    count_contribs = n()
  )|>
  arrange(count_contribs)
```

**Q4** What's your best educated guess to explain why the city identified in question 2 had more contributions compared to the cities identified in question 3? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.

**A4**
Charleston is the capital and largest city in West Virginia, in comparison to Allen Junction which appears to be a small rural city in West Virginia. That being said the reason why there are more contributions for Charleston than Allen Junction probably has to due with both the the difference in population between those cities and the wealth of the people living in those cities. There are a lot more people living in Charleston and it is more of a developed city so the people in Charleston might be more able to give monitary contributions than Allen Junction a small rural, farm town in West Virginia

**Q5** What recipient got the largest total amount of money? How many contributions did it receive? And what was the average donation it received?

Hint: Since there three distinct questions here, you'll need to do multiple things inside summarise() this time.
Hint 2: "aggregate_amount" is NOT the column you want to use here.
**A5**
TRUMP NATIONAL COMMITTEE JFC, INC received the largest total amount of money, they received $463,944.40 over 14,063 contributions. The average donation was $33. 
```{r}
westvirginia_winred_contributions |>
 group_by(memo_text) |>
  summarise( 
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
  ) |>
  arrange(desc(total_amount))
```

**Q6** What is that top recipient and what explains its popularity? Hint: this will require web research, you won't be able to answer this purely from the data set. You do not need to write any code for this question, so no codeblock is provided.

**A6**
It is a joint fund, so when people donate to that group they aren't funding Trump or a specific republican candidate. They are donating money to and the Trump republican team is sending amounts of that money to whichever candidate their team feels needs the most help and finances.

**BONUS POINTS (do this last / only if you have time at the end)** you might notice that a lot of the top recipient committees seem to be tied to the same person, and some even have nearly-identical committee names. What might be going on here? (You don't need to get this RIGHT, just give me an educated guess or two.)

**A**
Winred is a republican platform that encourages republican voters to donate money to republican candidates, so it would make the most sense for a lot of the recipient committees to use Trumps name because he is the president from the republican party. It encourages those who support the party to send money to "Trump" committees by using his name as support

**Q7**. Using the answer from A5 as a starting point, produce a summary table that has the following information for each recipient: \* count of contributions \* total amount of all contributions \* the average contribution amount \* the median contribution amount \* the smallest contribution amount \* the largest contribution amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7**.
Something that I saw that might be newsworthy was that Carol for Congress recipient had the 9th most total donations but when looking at average of the donations, had the highest average donation of $1490, out of 15 contribution to her. I am not counting the fair counts courts and Guthrie for Congress since they only received one total contribution.
```{r}
wv_summary <- westvirginia_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) |>
  arrange(desc(total_amount))
```
